% ============================================================
% Section 2: Related Work
% ============================================================
\section{Related Work}

\subsection{Deep Learning for Dental Fluorosis Diagnosis}

Automated dental fluorosis (DF) diagnosis using deep learning has seen several recent advances. Xu et al. proposed MLTrMR \cite{xu2025mltrmr}, a Masked Latent Transformer with random masking ratio, achieving 80.19\% accuracy on a 200-image balanced 4-class DF dataset (DFID). The architecture uses a 556M-parameter latent transformer with multi-stage training. LD2Net \cite{li2026ld2net} proposed a lightweight dual-network architecture achieving 80.00\% accuracy with only 3.31M parameters, targeting edge deployment. Other methods including FusionDentNet and HiFuse have reported results in the $\sim$78--80\% accuracy range on the same benchmark.

All existing approaches share a common paradigm: nominal classification via cross-entropy loss on majority-vote one-hot labels. This ignores three clinically relevant aspects: (1) the ordinal nature of fluorosis grades (Normal $<$ Mild $<$ Moderate $<$ Severe), where adjacent-grade errors carry different clinical weight than distant-grade errors; (2) prediction uncertainty, which is critical for knowing when to trust or defer an automated diagnosis; and (3) the subjective nature of visual fluorosis grading, where different clinicians may assign different grades to the same image. Our work addresses all three gaps through the EDL+ORCU framework.

\subsection{Evidential Deep Learning}

Evidential Deep Learning (EDL), introduced by Sensoy et al. \cite{sensoy2018edl}, replaces softmax outputs with Dirichlet distribution parameters. The model outputs concentration parameters $\alpha_k > 1$, from which class probabilities $p_k = \alpha_k / S$, evidence $e_k = \alpha_k - 1$, and predictive uncertainty $u = K / S$ are derived, where $S = \sum_k \alpha_k$. Training minimizes Type II Maximum Likelihood with a KL divergence regularizer that shrinks non-target evidence toward unity, encouraging the model to accumulate evidence only for the predicted class.

EDL has been successfully applied to skin lesion classification, histopathology image analysis, and COVID-19 diagnosis from chest X-rays, demonstrating that Dirichlet-based uncertainty is more informative for clinical decisions than entropy-based or ensemble-based alternatives \cite{ulmer2022prior}. However, EDL has not been previously applied to fluorosis diagnosis, nor has it been combined with ordinal regression constraints in a shared-logit architecture. Our work is the first to demonstrate EDL's effectiveness for dental fluorosis grading, including its regularization benefits for small-sample initialization robustness.

\subsection{Ordinal Regression and Calibration}

Ordinal regression addresses classification with ordered labels, where the cost of misclassification depends on the distance between predicted and true classes. Standard deep learning approaches include threshold-based cumulative link models and loss modifications that penalize order violations.

Kim et al. recently proposed ORCU \cite{kim2025orcu}, combining SORD (Soft-Encoded Ordinal Regression) with log-barrier regularization that enforces logit monotonicity. SORD replaces one-hot labels with soft target distributions based on squared ordinal distance: $y^{\text{SORD}}_k = \text{softmax}(-(k - y)^2)$. The log-barrier penalty on adjacent logit differences $r_k = z_k - z_{k+1}$ ensures that predicted probability distributions are unimodal---a desirable property for clinical interpretability, as multi-modal distributions (e.g., simultaneous high probability for both Normal and Severe) are clinically implausible. ORCU was originally demonstrated on age estimation and depth prediction tasks. We adapt ORCU to medical ordinal classification through the novel shared-logit architecture, where the same logits $z$ feeding the ORCU regularization also produce the Dirichlet parameters for EDL.

\subsection{Calibration and Uncertainty in Medical Imaging}

Model calibration---the alignment between predicted confidence and empirical accuracy---is essential for clinical deployment. Guo et al. \cite{guo2017calibration} demonstrated that modern deep neural networks are systematically miscalibrated, typically overestimating confidence. Standard mitigation approaches include temperature scaling (a single parameter $T > 0$ applied post-hoc to logits) and ensemble methods \cite{lakshminarayanan2017ensemble}, but these provide only confidence calibration without distinguishing between aleatoric (data-inherent) and epistemic (model-knowledge) uncertainty.

EDL offers a principled alternative: the Dirichlet distribution naturally separates total uncertainty $u$ into evidence per class, with low evidence across all classes indicating high uncertainty. This is particularly relevant for small-sample medical imaging, where epistemic uncertainty dominates and standard calibration methods may be insufficient.
