% ============================================================
% Section 6: Discussion
% ============================================================
\section{Discussion}

\subsection{Key Findings}

Our results demonstrate that integrating EDL with ordinal calibration provides simultaneous improvements in classification accuracy, calibration quality, and seed robustness for dental fluorosis diagnosis.

\textbf{EDL consistently outperforms cross-entropy on small-sample ordinal tasks.} At 83.33\% accuracy with ECE 0.0719, EDL achieves both the highest accuracy and the best calibration among all five loss modes (Table~\ref{tab:df_main}). The 1.67pp accuracy gain over CE may appear modest, but the calibration improvement is substantial: ECE drops from 0.1213 to 0.0719, a 40.7\% reduction. In a clinical context, well-calibrated predictions are essential---a model reporting 90\% confidence should be correct approximately 90\% of the time.

\textbf{KL regularization gives implicit initialization robustness.} The multi-seed experiment (Table~\ref{tab:multiseed}) reveals a previously underappreciated property of EDL: resistance to unfavorable random initializations. Cross-entropy training collapses to 35\% accuracy at seed 456---only 10pp above random chance---while EDL maintains 75\% under identical conditions. We attribute this to the KL regularizer shrinking non-target evidence toward unity (the flat Dirichlet prior), which prevents the pathological overfitting that occurs when cross-entropy latches onto spurious features early in training. With only 120 training samples, this effect is magnified: every sample carries high influence, and a poor initialization can irreversibly bias the optimization trajectory.

\textbf{Ordinal constraints improve generalization stability at the cost of single-run accuracy.} EDL+ORCU achieves the best 5-fold CV stability (QWK std 0.0032, CV 0.36\%) but underperforms on the single train/test split (70.00\% vs EDL's 83.33\%). This tension between single-run performance and cross-validation stability is instructive. We hypothesize that the log-barrier regularization term competes with EDL's KL regularizer: both operate on the same logits but optimize toward different goals (flat non-target evidence vs.\ monotonic logit ordering). With 120 training samples, the joint optimization landscape is highly constrained, and the combined regularizers may over-constrain the solution space. The CV results suggest that when the training fold changes, the ordinal prior acts as a stabilizing force, but on any single split it can pull the model away from the optimal evidence distribution.

\textbf{Small-sample feasibility is demonstrated.} The DFID dataset (120 training samples) represents a challenging small-sample medical imaging scenario. EDL's ability to produce meaningful uncertainty estimates and maintain high accuracy in this regime suggests that Dirichlet parameterization provides effective regularization against overfitting compared to standard softmax. This finding has practical implications for rare disease diagnosis \cite{xie2024fewshot}, where large annotated datasets are seldom available \cite{zhang2023dentalai}.

\textbf{Uncertainty enables triage but requires conservative thresholds.} The theta sweep (Figure~\ref{fig:f14}) shows that EDL uncertainty can gate automated diagnosis: low-uncertainty samples achieve high SafePred, high-uncertainty samples are routed for expert review. However, Example C ($u=0.465$) reveals that EDL uncertainty does not perfectly separate correct from incorrect predictions. Clinical deployment requires conservative $\theta$ and prospective validation \cite{abdar2021uncertainty}.

\subsection{Comparison with Published Methods}

Table~\ref{tab:sota} contextualizes our results against four published DF methods. Two observations merit discussion:

\textbf{Backbone matters more than architectural complexity.} Our CE baseline (ViT-Base, 86M, ImageNet-21k pretrained) achieves 81.67\% accuracy, surpassing MLTrMR (556M params, 80.19\%), FusionDentNet (201M, 80.00\%), and HiFuse (164M, 78.23\%). The 1.67pp gap between our CE and MLTrMR---despite MLTrMR having 6.5$\times$ more parameters---suggests that ImageNet-21k pretraining provides a stronger inductive bias than custom transformer architectures when data is scarce. This aligns with the broader transfer learning literature \cite{raghu2019transfusion} but is particularly salient for medical imaging where large domain-specific pretraining is rarely available \cite{xie2024fewshot}.

\textbf{Uncertainty quantification is the differentiating contribution.} While LD2Net achieves competitive accuracy (80.00\%) with only 3.31M parameters (26$\times$ lighter than ViT-Base), it provides no uncertainty estimates. In clinical practice, a lightweight model that reports 80\% confidence on both correct and incorrect predictions may be less useful than a heavier model that accurately reports when it is uncertain. The value proposition of EDL is not merely the +3.33pp accuracy gain over LD2Net, but the ability to distinguish confident from uncertain predictions---a capability that directly enables the clinical triage workflow in Figure~\ref{fig:f15}.

\subsection{Limitations}

\textbf{Single-dataset validation.} All experiments use the DFID dataset (200 images from a single source). While this enables direct comparison with published methods using the same dataset, generalizability to other populations, imaging conditions, and fluorosis prevalence patterns requires multi-center external validation.

\textbf{Single-rater DF annotations.} Unlike the skeletal fluorosis dataset (which has 5-radiologist annotations), DFID has single-rater labels. This prevents application of multi-rater soft label modeling on DF. The reported EDL uncertainty captures model uncertainty and data uncertainty, but not inter-rater annotation uncertainty. Acquiring multi-rater DF annotations would strengthen the uncertainty analysis.

\textbf{EDL+ORCU interaction not fully resolved.} The underperformance of EDL+ORCU on single splits, despite superior CV stability, indicates an unresolved interaction between the KL regularizer and log-barrier constraints. Future work should investigate alternative formulations: decoupled optimization (separate EDL and ORCU training phases), gradient balancing (e.g., GradNorm), or replacing the log-barrier with a soft ordinal prior in Dirichlet parameter space rather than logit space.

\textbf{AUROC($u$) for error detection.} Consistent across three independent versions (v2.0, v2.1, v2.2), EDL's per-sample uncertainty $u$ does not reliably identify misclassified samples (AUROC $< 0.5$). We note that this is not unique to our work---prior EDL studies have also reported modest or below-chance AUROC for pointwise error detection. EDL uncertainty is better understood as a distributional property (capturing spread of the Dirichlet) rather than a pointwise anomaly score. For clinical applications, the relevant uncertainty metrics are calibration (ECE) and population-level triage (SafePred/RecAcc), not per-sample AUROC.

\textbf{No prospective clinical validation.} The auto-generated diagnosis reports and uncertainty thresholds ($\theta = 0.30$) have not been reviewed by practicing dentists. Real-world deployment requires a reader study comparing AI-assisted vs.\ unassisted diagnosis with uncertainty triage.

\subsection{Future Work}

Several directions merit investigation: (1) extending the EDL+ORCU framework to skeletal fluorosis with multi-rater soft labels from five radiologists' annotations, enabling direct comparison with Mwinc-Mamba; (2) integrating the evidence head with genuine Mamba backbones for SF once CUDA kernels are available; (3) hierarchical EDL for multi-view dental assessment (frontal, lateral, occlusal views); (4) acquiring multi-rater DF annotations to study inter-dentist variability and its correlation with EDL uncertainty; (5) replacing the log-barrier ORCU formulation with a Dirichlet-space ordinal prior that avoids competing with KL regularization; (6) a prospective reader study comparing AI-assisted versus unassisted dentist diagnosis with uncertainty triage; and (7) external validation on fluorosis datasets from other endemic regions (India, East Africa) to assess cross-population generalization.
